---
title: "Analysis1"
output: html_document
date: "2024-04-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(tidyr)
```

```{r}
corn_df = read_csv("crops.csv")
```

```{r}
bushels_df = corn_df %>%
  select("County", "Ag District", "Year", "CORN, GRAIN - PRODUCTION, MEASURED IN BU", "SOYBEANS - PRODUCTION, MEASURED IN BU") %>%
  rename(total_corn = "CORN, GRAIN - PRODUCTION, MEASURED IN BU", district = "Ag District", total_soybeans = "SOYBEANS - PRODUCTION, MEASURED IN BU")
```

# Missing County Information

```{r}
bushels_df %>%
  group_by(Year) %>%
  summarize(num_na = sum(is.na(total_corn)), total=n(), prop_missing = num_na/total)
```

- This may effect the total number of corn/soybean bushels recorded
- Some counties may be missing. Either it was not recorded or maybe that county did not harvest that type of crop

```{r}
replace_na_with_mean_or_zero <- function(x) {
  if (any(!is.na(x))) {
    x[is.na(x)] <- mean(x, na.rm = TRUE)
    return(x)
  } else {
    x[is.na(x)] <- 0
    return(x)
  }
}

crops_df = read_csv("crops.csv") %>%
  rename(
    sugar_harvest = "SUGARBEETS - ACRES HARVESTED",
    sugar_planted = "SUGARBEETS - ACRES PLANTED",
    sugar_produced = "SUGARBEETS - PRODUCTION, MEASURED IN TONS",
    sugar_sucrose = "SUGARBEETS - SUCROSE, MEASURED IN PCT",
    sugar_per_acre = "SUGARBEETS - YIELD, MEASURED IN TONS / ACRE"
  ) %>%
  group_by(County) %>%
  mutate(sugar_harvest = replace_na_with_mean_or_zero(sugar_harvest),
         sugar_planted = replace_na_with_mean_or_zero(sugar_planted),
         sugar_produced = replace_na_with_mean_or_zero(sugar_produced),
         sugar_sucrose = replace_na_with_mean_or_zero(sugar_sucrose),
         sugar_per_acre = replace_na_with_mean_or_zero(sugar_per_acre),
         ) %>%
  rename(
    corn_harvested = "CORN, GRAIN - ACRES HARVESTED",
    corn_produced = "CORN, GRAIN - PRODUCTION, MEASURED IN BU",
    corn_grain_per_acre = "CORN, GRAIN - YIELD, MEASURED IN BU / ACRE",
    soybeans_harvested = "SOYBEANS - ACRES HARVESTED", 
    soybeans_produced = "SOYBEANS - PRODUCTION, MEASURED IN BU",
    soybeans_grain_per_acre = "SOYBEANS - YIELD, MEASURED IN BU / ACRE") %>%
  mutate(
    across(where(is.numeric), ~replace_na(., 0))
  ) %>%
  mutate(
    soybeans_grain_per_acre = soybeans_grain_per_acre * 30.027682980816,
    corn_grain_per_acre = corn_grain_per_acre * 30.027682980816,
    corn_produced = corn_produced * 56 / 2000,
    soybeans_produced = soybeans_produced * 60 /2000
  )

```

# Yearly Corn and SoyBean Total Production Trends

```{r}
state_corn_bushels_df <- crops_df %>%
  group_by(Year) %>%
  filter(!is.na(corn_produced)) %>%
  summarize(corn = sum(corn_produced) / 1000) 

state_soybeans_bushels_df <- crops_df %>%
  group_by(Year) %>%
  filter(!is.na(soybeans_produced)) %>%
  summarize(soybeans = sum(soybeans_produced) / 1000) 

state_sugar_bushels_df <- crops_df %>%
  group_by(Year) %>%
  filter(!is.na(sugar_produced)) %>%
  summarize(sugarbeets = sum(sugar_produced) / 1000)  # Corrected column name

combined_df <- merge(state_corn_bushels_df, state_soybeans_bushels_df, by = "Year", all = TRUE)
combined_df <- merge(combined_df, state_sugar_bushels_df, by = "Year", all = TRUE)

combined_df_long <- tidyr::pivot_longer(combined_df, 
                                        cols = c("corn", "soybeans", "sugarbeets"), 
                                        names_to = "Crop", 
                                        values_to = "Production")

ggplot(combined_df_long, aes(x = Year, y = Production, color = Crop)) +
  geom_line(size = 1) +  
  geom_point(size = 2) +  
  labs(x = "Year", y = "Tons Produced (Thousands)",  
       title = "Total Tons of Crop Produced Over Time") +
  scale_color_manual(values = c("corn" = "#FF5733", "soybeans" = "#33A1FF", "sugarbeets" = "green"), 
                     limits = c("corn", "soybeans", "sugarbeets")) +  # Corrected color values
  theme_minimal() +  
  theme(
    plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),  
    axis.title = element_text(size = 14, face = "bold"),  
    axis.text = element_text(size = 12),  
    legend.title = element_text(size = 14, face = "bold"),  
    legend.text = element_text(size = 12) 
  )
```

```{r}
cor(state_soybeans_bushels_df$soybeans, state_corn_bushels_df$corn)
```

Observations:
- A lot more corn is produced than soybeans
- Both corn and soybeans seem to be following a similar trend(increase and decrease at similar times). This suggests that weather and environmental factors may influence both crops.
- There is a correlation of 83.2% which means that the total production of both is strongly correlated.

# Crop per Acre Analysis

```{r}
crop_df <- corn_df %>%
  rename(
    corn_harvested = "CORN, GRAIN - ACRES HARVESTED",
    corn_produced = "CORN, GRAIN - PRODUCTION, MEASURED IN BU",
    corn_grain_per_acre = "CORN, GRAIN - YIELD, MEASURED IN BU / ACRE",
    soybeans_harvested = "SOYBEANS - ACRES HARVESTED", 
    soybeans_produced = "SOYBEANS - PRODUCTION, MEASURED IN BU",
    soybeans_grain_per_acre = "SOYBEANS - YIELD, MEASURED IN BU / ACRE") %>%
  select(Year, corn_harvested, corn_produced, corn_grain_per_acre, soybeans_harvested, soybeans_produced, soybeans_grain_per_acre) %>%
  mutate(
    across(where(is.numeric), ~replace_na(., 0))
  )
```

```{r}
grain_per_acre_df = crop_df %>%
  group_by(Year) %>%
  summarize(
    total_corn_harvested = sum(corn_harvested),
    total_corn_produced = sum(corn_produced),
    total_soybeans_harvested = sum(soybeans_harvested),
    total_soybeans_produced = sum(soybeans_produced)
  ) %>%
  mutate(
    corn_per_acre = total_corn_produced/total_corn_harvested,
    soybeans_per_acre = total_soybeans_produced/total_corn_harvested
  )
```

```{r}
ggplot(grain_per_acre_df, aes(x = Year)) +
  geom_line(aes(y = corn_per_acre, color = "Corn")) +
  geom_line(aes(y = soybeans_per_acre, color = "Soybeans")) +
  labs(x = "Year", y = "Bushels per Acre", color = "Crop", title = "Corn and Soybean Bushel Production per Acre Over Years") +
  scale_color_manual(values = c("Corn" = "blue", "Soybeans" = "red"))
```

Observation:
- We can grow more corn per acre than we can grow soybeans per acre
- The trends of crops per acre seem strongly correlated

```{r}
cor(grain_per_acre_df$corn_per_acre, grain_per_acre_df$soybeans_per_acre)
```

- They are strongly correlated

# Acre Usage for each type of crop over the years
```{r}
ggplot(grain_per_acre_df, aes(x = Year)) +
  geom_line(aes(y = total_corn_harvested, color = "Corn")) +
  geom_line(aes(y = total_soybeans_harvested, color = "Soybeans")) +
  labs(x = "Year", y = "Acres Harvested", color = "Crop", title = "Total Acres Harvested for Corn and Soybeans Over Time") +
  scale_color_manual(values = c("Corn" = "blue", "Soybeans" = "red"))
```

```{r}
grain_per_acre_df = grain_per_acre_df %>%
  mutate(total_harvested = total_corn_harvested + total_soybeans_harvested)

ggplot(grain_per_acre_df, aes(x = Year)) +
  geom_line(aes(y = total_harvested)) +
  labs(x = "Year", y = "Acres Harvested", title = "Total Acres Harvested Over Time")
```

TODO: Visualizing the changes in Total and Per Acre to better understand that they are strongly correlated.
